## 2.3 Lab: Introduction to R

### 2.3.1 Basic Commands

Here's an example of the R function [`c()`](http://bit.ly/R_c) that we can use to concatenate values into a vector or a list.

```{r}
x <- c(1, 3, 2, 5)
x
```


```{r}
x <- c(1, 6, 2)
x
y <- c(1, 4, 3)
```

We can check the length of a vector or list with the [`length()`](http://bit.ly/R_length) function.

```{r}
length(x)
length(y)
x + y
```

To get a list of all objects in the current environment we use the [`ls()`](http://bit.ly/R_ls) function. Objects can be removed from the environment with the [`rm()`](http://bit.ly/R_rm) function.

```{r}
ls()
rm(x, y)
ls()
```

To remove all objects from the environment, we first get the list of all objects with the [`ls()`](http://bit.ly/R_ls) function, and pass this list to the [`rm()`](http://bit.ly/R_rm) function. 

```{r}
rm(list = ls())
```

To get help on any function, we can use the `?` operator.
```{r eval=FALSE}
?matrix
```

A matrix is defined using the [`matrix()`](http://bit.ly/R_matrix) function.

```{r}
x=matrix(data=c(1,2,3,4), nrow=2, ncol=2)
x
```

When calling a function in R, positional arguments can be speficied without the names of the arguments as shown below. Note that the resulting matrix is identical to the one in the previous call to [`matrix()`](http://bit.ly/R_matrix)

```{r}
x <- matrix(c(1, 2, 3, 4), 2, 2)
x
```

Finally, we can also tell the [`matrix()`](http://bit.ly/R_matrix) function whether the data provided are in row-order or column-order. The following example shows, matrix creation with data in row-order.

```{r}
matrix(c(1, 2, 3, 4), 2, 2, byrow = TRUE)
```

We can take the square-root of each element of the matrix with the [`sqrt()`](http://bit.ly/R_math_fun) function and use the `^` operator raise each element of the matrix to a power.

```{r}
sqrt(x)
x^2
```

The [`rnorm()`](http://bit.ly/R_normal) function can be used to generate random numbers. In the following example, we use [`rnorm()`](http://bit.ly/R_normal) to generate a vector of `50` random numbers.

```{r}
x <- rnorm(50)
y <- x + rnorm(50, mean = 50, sd = 0.1)
cor(x, y)
```

To get consistent results from repeated calls to [`rnorm()`](http://bit.ly/R_normal), we use the [`set.seed()`](http://bit.ly/R_random). 

```{r}
set.seed(1303)
rnorm(50)
```

The functions [`mean()`](http://bit.ly/R_mean), [`var()`](http://bit.ly/R_cor), and [`sd()`](http://bit.ly/R_sd) can be used to calculate the mean, variance and standard deviation of a vector.

```{r}
set.seed(3)
y <- rnorm(100)
mean(y)
var(y)
sqrt(var(y))
sd(y)
```

### 2.3.2 Graphics

Basic plotting functionality is provided by the [`plot()`](http://bit.ly/R_plot) function in R. The following example shows how to produce a scatter plot and how to label the x and y axis.

```{r}
x <- rnorm(100)
y <- rnorm(100)
plot(x, y)
plot(x, y, xlab = "this is the x-axis", ylab = "this is the y-axis", main = "Plot of X vs Y")
```

We can save the plots generated in R to a file in PDF format, or to a graphics image format such as JPEG or PNG. In the following example, we use the [`pdf()`](http://bit.ly/R_pdf) function to create a PDF file called `Figure.pdf`. 

```{r}
pdf("Figure.pdf")
plot(x, y, col = "green")
dev.off()
```

To generate sequence of numbers, we can call the [`seq()`](http://bit.ly/R_seq) function.

```{r}
x <- seq(1, 10) > x
x <- 1:10
x
x <- seq(-pi, pi, length = 50)
```

Contour plots can be created in R with the [`contour()`](http://bit.ly/R_contour) function.

```{r}
y <- x
f <- outer(x, y, function(x, y) cos(y)/(1 + x^2))
contour(x, y, f)
contour(x, y, f, nlevels = 45, add = T)
fa <- (f - t(f))/2
contour(x, y, fa, nlevels = 15)
```

Heatmaps can be generated with the [`image()`](http://bit.ly/R_image) function.

```{r}
image(x, y, fa)
```

Perspective plots of a three-dimentional surface are plottted using the [`persp()`](http://bit.ly/R_persp) function.

```{r}
persp(x, y, fa)
persp(x, y, fa, theta = 30)
persp(x, y, fa, theta = 30, phi = 20)
persp(x, y, fa, theta = 30, phi = 70)
persp(x, y, fa, theta = 30, phi = 40)
```

### 2.3.3 Indexing Data

Indexing elements of a vector or a matrix is done using the square-brackets `[` and `]`. In this example we create a 4x4 matrix and then index the element at row 2, column 3.

```{r}
A <- matrix(1:16, 4, 4)
A
```

```{r}
A[2, 3]
```

We can also select multiple rows or columns at once by providing the row or the column indices as vectors instead of single integers.

```{r}
A[c(1, 3), c(2, 4)]
A[1:3, 2:4]
A[1:2, ]
A[, 1:2]
```

If a row or column index is omitted, then R assumes that we're looking for all rows or all columns.

```{r}
A[1, ]
```

To exclude certain elements, we can use the negative sign `-` before the index.

```{r}
A[-c(1, 3), ]
A[-c(1, 3), -c(1, 3, 4)]
```

The [`dim()`](http://bit.ly/R_dim) can be used to show the dimensions of a vector or a matrix.

```{r}
dim(A)
```

### 2.3.4 Loading Data

The [`read.table()`](http://bit.ly/R_read_table) and  [`read.csv()`](http://bit.ly/R_read_table) functions are used to import data from a file into R.

```{r}
Auto <- read.table("http://www-bcf.usc.edu/~gareth/ISL/Auto.data")
head(Auto)
```

The [`read.table()`](http://bit.ly/R_read_table) and  [`read.csv()`](http://bit.ly/R_read_table) family of functions offer a number of options for controlling how the data are imported, such as control over header rows, blank rows, or missing values.

```{r}
Auto <- read.csv("http://www-bcf.usc.edu/~gareth/ISL/Auto.csv", header = T, na.strings = "?")
head(Auto)
dim(Auto)
Auto[1:4, ]
```

```{r}
Auto <- na.omit(Auto)
dim(Auto)
```

We can get a list of all variables in the dataset we imported with the [`names()`](http://bit.ly/R_names) function.

```{r}
names(Auto)
```

### 2.3.5 Additional Graphical and Numerical Summaries

There are two ways to reference a variable from a dataset. The first uses the `$` symbol to tell R that the variable `cylinders` is from the `Auto` dataset.

```{r}
plot(Auto$cylinders, Auto$mpg)
```

The second method uses the [`attach()`](http://bit.ly/R_attach) function which tells R to search for variable names in the `Auto` dataset.

```{r}
attach(Auto)
plot(cylinders, mpg)
```

We can convert quantitative variables to qualitative or categorical variables with the  [`as.factor()`](http://bit.ly/R_factor) function.

```{r}
cylinders <- as.factor(cylinders)
```

Catrgotrical variables are automatically plotted as box-plots by R.

```{r}
plot(cylinders, mpg)
plot(cylinders, mpg, col = "red")
plot(cylinders, mpg, col = "red", varwidth = T)
plot(cylinders, mpg, col = "red", varwidth = T, horizontal = T)
plot(cylinders, mpg, col = "red", varwidth = T, xlab = "cylinders", ylab = " MPG ")
```

We can also get a hsitogram of the data with the  [`hist()`](http://bit.ly/R_hist) function.

```{r}
hist(mpg)
hist(mpg, col = 2)
hist(mpg, col = 2, breaks = 15)
```

To generate a matrix of scatterplots, we can use the  [`pairs()`](http://bit.ly/R_pairs) 

```{r}
pairs(Auto)
pairs(~mpg + displacement + horsepower + weight + acceleration, Auto)
```

The [`identify()`](http://bit.ly/R_identify) function can be used to identify specific points on a plot. In an interactive session in R-Studio, [`identify()`](http://bit.ly/R_identify) allows you to click on specific points and labels them on screen for convenient identification.

```{r}
plot(horsepower, mpg)
identify(horsepower, mpg, name)
```

The  [`summary()`](http://bit.ly/R_summary) function can be used to print a summary of the dataset.

```{r}
summary(Auto)
```

```{r}
summary(mpg)
```


