## Installation

In order to build the book and publish it on GitHub, you need the following:

* Node.js Package Manager (NPM)
    * Download and install it from https://nodejs.org

* Gitbook
    * On Linux and Mac OSX systems, you can install Gitbook from the command line through `npm`. Since `npm` install requires administrative priviliges, make sure you have the correct access level or use `sudo`.
    
        ```bash
        npm install gitbook -g
        npm install gitbook-cli -g
        ```
        
* RGitbook
    * If you don't have `devtools` package installed, then install it first
        ```{r eval=FALSE}
        install.packages("devtools")
        ```

    * Once `devtools` is installed, get the following fork of `RGitbook` from GitHub.
    
        ```{r eval=FALSE}
        devtools::install_github("altaf-ali/RGitbook")
        ```
    
## Building

You can build the book with the buildGitbook() function from the `Rgitbook` package.

```{r eval=FALSE}
Rgitbook::buildGitbook()
```

## Cleanup

If the build process fails or you get unexpected results, then you might want to clean the build area first before running the build.

```{r eval=FALSE}
Rgitbook::cleanGitbook()
```

## Testing

Test the gitbook before publishing it on GitHub. 

```{r eval=FALSE}
Rgitbook::openGitbook()
```

## Publishing

Once the book is built, you can publish it with `publishGitbook()`

```{r eval=FALSE}
Rgitbook::publishGitbook()
```

