## 3.6 Lab: Linear Regression

### 3.6.1 Libraries

We use [`library()`](http://bit.ly/R_library) to access functionality provided by packages not included in the standard R installation. To install a new package, use the [`install.packages()`](http://bit.ly/R_install_packages) function from the command line console.

```{r eval=FALSE}
install.packages("ISLR")
```

Start by loading that `MASS` and `ISLR` packages that we will be using throughtout this exercise

```{r}
library(MASS)
library(ISLR)
```

### 3.6.2 Simple Linear Regression

To access the `Boston` housing dataset from the `MASS` package, use the [`attach()`](http://bit.ly/R_attach) function

```{r, message=FALSE, warning=FALSE}
attach(Boston)
```

We can use [`head()`](http://bit.ly/R_head) and [`names()`](http://bit.ly/R_names) to take a look at the Boston dataset

```{r}
head(Boston)
names(Boston)
```

Now, lets start with a simple linear model with [`lm()`](http://bit.ly/R_lm)

```{r}
lm.fit <- lm(medv ~ lstat)
```

Instead of attaching the `Boston` dataset first, we can also specify it directly in [`lm()`](http://bit.ly/R_lm).

```{r}
lm.fit <- lm(medv ~ lstat, data = Boston)
```

We can examine the fitted model, by simply typing its name.

```{r}
lm.fit
```

The [`names()`](http://bit.ly/R_names) function lists the names of all variables store in the `lm.fit` model.

```{r}
names(lm.fit)
```

We can extract the coefficients estimated by [`lm()`](http://bit.ly/R_lm) using the [`coef()`](http://bit.ly/R_coef) function, and the confidence interval with the [`confint()`](http://bit.ly/R_coef) function.

```{r}
coef(lm.fit)
```

```{r}
confint(lm.fit)
```

We use the [`predict()`](http://bit.ly/R_predict) function to obtain prediction interval or confidence intervals for a given value of `lstat`

```{r}
predict(lm.fit, data.frame(lstat = (c(5, 10, 15))), interval = "confidence")
```

```{r}
predict(lm.fit, data.frame(lstat = (c(5, 10, 15))), interval = "prediction")
```

We can plot the variables `lstat` and `medv` using the [`plot()`](http://bit.ly/R_plot) function, and overlay a regression line from our linear model using the [`abline()`](http://bit.ly/R_abline) function.

```{r}
plot(lstat, medv)
abline(lm.fit)
```

Experiment with different options for [`abline()`](http://bit.ly/R_abline) by changing the line width and color.

```{r}
plot(lstat, medv)
abline(lm.fit, lwd = 3)
```

```{r}
plot(lstat, medv)
abline(lm.fit, lwd = 3, col = "red")
```

```{r}
plot(lstat, medv, col = "red")
plot(lstat, medv, pch = 20)
plot(lstat, medv, pch = "+")
plot(1:20, 1:20, pch = 1:20)
```

To create a grid of multiple subplots, use the [`par()`](http://bit.ly/R_par) function.

```{r}
par(mfrow = c(2, 2))
plot(lm.fit)
```

We can extract the residuals and studentized residuals from our linear model using [`residuals()`](http://bit.ly/R_residuals) and [`rstudent()`](http://bit.ly/R_influence_measures), and plot them along with the predicted values.

```{r}
plot(predict(lm.fit), residuals(lm.fit))
plot(predict(lm.fit), rstudent(lm.fit))
```

We can compute the influence matrix for the predictors in our fitted model using the [`hatvalues()`](http://bit.ly/R_influence_measures) function

```{r}
plot(hatvalues(lm.fit))
which.max(hatvalues(lm.fit))
```

### 3.6.3 Multiple Linear Regression

The [`lm()`](http://bit.ly/R_lm) can be used for fitting multiple regression models as well. In the following example, we include `age` in addition to `lstata` as predictors of of our response variable `medv`. 

```{r}
lm.fit <- lm(medv ~ lstat + age, data = Boston)
summary(lm.fit)
```

In a formula, a dot `.` can be used as a shorthand to include all variables from the `Boston` dataset as predictors.

```{r}
lm.fit <- lm(medv ~ ., data = Boston)
summary(lm.fit)
```

The Variance Inflation Factors (VIF) can be caculated using the `vif()` function from the `Companion to Applied Regression` or `car` package. You can install the `car` package using [`install.packages()`](http://bit.ly/R_install_packages) if it is not already installed.

```{r eval=FALSE}
install.packages("ISLR")
```

```{r}
library(car)
vif(lm.fit)
```

To exclude specific variables from the list of predictors, we can use the `-` notation. In the following example, all variables except `age` are included in the linear fit model.

```{r}
lm.fit1 <- lm(medv ~ . - age, data = Boston)
summary(lm.fit1)
```

We can also update an existing fit with the [`update()`](http://bit.ly/R_update) function and specifiying a new formula.

```{r}
lm.fit1 <- update(lm.fit, ~. - age)
```

### 3.6.4 Interaction Terms

The `:` and `*` operators are used to specify interaction terms between variables. The `:` operator includes just the interaction term between the specified variables as predictor, while the `*` operator includes the interaction term as well as the variables themselves. In the following example, `lstat * age` is equivalent to `lstat + age + lstat:age`

```{r}
summary(lm(medv ~ lstat * age, data = Boston))
```

### 3.6.5 Non-linear Transformations of the Predictors

The [`lm()`](http://bit.ly/R_lm) function can also be non-linear transformations such as quadratic or cubic terms. Terms in the formula raised to a power with `^` must be wrapped inside a call to [`I()`](http://bit.ly/R_asis) to prevent interpretation and treated `as is`.

```{r}
lm.fit2 <- lm(medv ~ lstat + I(lstat^2))
summary(lm.fit2)
```

To examine the analysis of variance (ANOVA) for one or more models, we use the [`anova()`](http://bit.ly/R_anova).

```{r}
lm.fit <- lm(medv ~ lstat)
anova(lm.fit, lm.fit2)
```

```{r}
par(mfrow = c(2, 2))
plot(lm.fit2)
```

```{r}
lm.fit5 <- lm(medv ~ poly(lstat, 5))
summary(lm.fit5)
```

In addition to polynomial transformations, the formula in a linear fit model can also include [`anova()`](http://bit.ly/R_anova) transformations as shown in the following example.

```{r}
summary(lm(medv ~ log(rm), data = Boston))
```

### 3.6.6 Qualitative Predictors

In this section, we will use the `Carseats` dataset from the `ISLR` package. Similar to how we used the `Boston` dataset, we can make the `Carseats` dataset available to us with the [`attach()`](http://bit.ly/R_attach) function.

```{r, message=FALSE, warning=FALSE}
attach(Carseats, warn.conflicts = FALSE)
```

```{r}
head(Carseats)
names(Carseats)
```

When dealing with qualitative or categorial variable, `R` automatically generates dummy variables for us. 

```{r}
lm.fit <- lm(Sales ~ . + Income:Advertising + Price:Age, data = Carseats)
summary(lm.fit)
```

To examine the coding for the qualitive variables, we can use the [`contrasts()`](http://bit.ly/R_contrasts) function.

```{r}
contrasts(ShelveLoc)
```

### 3.6.7 Writing Functions

We can define our own functions to wrap a set of `R` commands in a single call. In the following example, we define a `LoadLibraries` function that loads the `ISLR` and `MASS` packages.

```{r}
LoadLibraries <- function() {
    library(ISLR)
    library(MASS)
}
```

The function then simply be called just like any other function supplied with `R`.

```{r}
LoadLibraries() 
```

