## 9.6 Lab: Support Vector Machines

### 9.6.1 Support Vector Classifier

In this exercise we will use a support vector classifier with a linear kernel. We start by generating observations from two classes using the [`rnorm()`](http://bit.ly/R_normal) function.

```{r}
set.seed(1)
x <- matrix(rnorm(20 * 2), ncol = 2)
y <- c(rep(-1, 10), rep(1, 10))
x[y == 1, ] <- x[y == 1, ] + 1
```

```{r}
plot(x, col = (3 - y))
```

The `e1071` package provides functionality to fit a Support Vector Machine classifier.

```{r}
dat <- data.frame(x = x, y = as.factor(y))
library(e1071)
svmfit <- svm(y ~ ., data = dat, kernel = "linear", cost = 10, scale = FALSE)
```

The fitted model from the [`svm()`](http://bit.ly/R_svm) function can be plotted with [`plot()`](http://bit.ly/R_plot).

```{r}
plot(svmfit, dat)
```

We can inspect the support vectors with the `index` component of the classifier returned from [`svm()`](http://bit.ly/R_svm).

```{r}
svmfit$index
```

The [`summary()`](http://bit.ly/R_summary) function can be used to print a summary of the classifier.

```{r}
summary(svmfit)
```

We can change the `cost` parameter and re-run the classifier to see how the support vectors change.

```{r}
svmfit <- svm(y ~ ., data = dat, kernel = "linear", cost = 0.1, scale = FALSE)
plot(svmfit, dat)
svmfit$index
```

We can run cross-validation on a range of values for the `cost` parameter using the [`tune()`](http://bit.ly/R_tune) function.

```{r}
set.seed(1)
tune.out <- tune(svm, y ~ ., data = dat, kernel = "linear", ranges = list(cost = c(0.001, 0.01, 0.1, 1, 5, 10, 100)))
```

```{r}
summary(tune.out)
```

The best performing model is stored in the `best.model` component of the object returned from [`tune()`](http://bit.ly/R_tune).

```{r}
bestmod <- tune.out$best.model
summary(bestmod)
```

We can also classify a different set of observations using the best model obtained from [`tune()`](http://bit.ly/R_tune). First we create a new set of observations with the  [`rnorm()`](http://bit.ly/R_normal) function.

```{r}
xtest <- matrix(rnorm(20 * 2), ncol = 2)
ytest <- sample(c(-1, 1), 20, rep = TRUE)
xtest[ytest == 1, ] <- xtest[ytest == 1, ] + 1
testdat <- data.frame(x = xtest, y = as.factor(ytest))
```

We can then use the [`predict()`](http://bit.ly/R_predict) function to classify the observations in `testdat`.

```{r}
ypred <- predict(bestmod, testdat)
table(predict = ypred, truth = testdat$y)
```

We can also try a different value for the `cost` parameter and train the classifier again.

```{r}
svmfit <- svm(y ~ ., data = dat, kernel = "linear", cost = 0.01, scale = FALSE)
ypred <- predict(svmfit, testdat)
table(predict = ypred, truth = testdat$y)
```

We can linearly separate the two classes and plot the observations.

```{r}
x[y == 1, ] <- x[y == 1, ] + 0.5
plot(x, col = (y + 5)/2, pch = 19)
```

We can fit a classifier with a large value for the `cost` parameter to see how the observations are classified.

```{r}
dat <- data.frame(x = x, y = as.factor(y))
svmfit <- svm(y ~ ., data = dat, kernel = "linear", cost = 1e+05)
summary(svmfit)
plot(svmfit, dat)
```

We can also try with a a smaller `cost` value.

```{r}
svmfit <- svm(y ~ ., data = dat, kernel = "linear", cost = 1)
summary(svmfit)
plot(svmfit, dat)
```

### 9.6.2 Support Vector Machine

We begin by generating data with non-linear class boundaries.

```{r}
set.seed(1)
x <- matrix(rnorm(200 * 2), ncol = 2)
x[1:100, ] <- x[1:100, ] + 2
x[101:150, ] <- x[101:150, ] - 2
y <- c(rep(1, 150), rep(2, 50))
dat <- data.frame(x = x, y = as.factor(y))
```

We can verify that the class boundaries are not linear by plotting the observations.

```{r}
plot(x, col = y)
```

We split the data into training and test subsets and run the SVM classifier with `kernel = "radial"` parameter.

```{r}
train <- sample(200, 100)
svmfit <- svm(y ~ ., data = dat[train, ], kernel = "radial", gamma = 1, cost = 1)
plot(svmfit, dat[train, ])
```

We can examine the model fit with the  [`summary()`](http://bit.ly/R_summary) function.

```{r}
summary(svmfit)
```

We can use a larger value for the `cost` parameter and see if it reduces the training errors.

```{r}
svmfit <- svm(y ~ ., data = dat[train, ], kernel = "radial", gamma = 1, cost = 1e+05)
plot(svmfit, dat[train, ])
```

We can run cross-validation using the  [`tune()`](http://bit.ly/R_tune) function.

```{r}
set.seed(1)
tune.out <- tune(svm, y ~ ., data = dat[train, ], kernel = "radial", ranges = list(cost = c(0.1, 1, 10, 100, 1000), gamma = c(0.5, 1, 2, 3, 4)))
summary(tune.out)
```

We can predict the classes on the test subset and examine the number of observations misclassified.

```{r}
table(true = dat[-train, "y"], pred = predict(tune.out$best.model, newdata = dat[-train, ]))
```

### 9.6.3 ROC Curves

We use the `ROCR` package to produce ROC curves on the predictions from the test subset.

```{r}
library(ROCR)
rocplot <- function(pred, truth, ...) {
    predob <- prediction(pred, truth)
    perf <- performance(predob, "tpr", "fpr")
    plot(perf, ...)
}
```

Instead of getting the class labels, we can also get fitted values from [`svm()`](http://bit.ly/R_svm) using `decision.values = TRUE` parameter.

```{r}
svmfit.opt <- svm(y ~ ., data = dat[train, ], kernel = "radial", gamma = 2, cost = 1, decision.values = TRUE)
fitted <- attributes(predict(svmfit.opt, dat[train, ], decision.values = TRUE))$decision.values
```

We generate the ROC curve with the [`rocplot()`](http://bit.ly/R_roc_plot) function. We can also change the value of $\gamma$ and see if it improves our predictions.

```{r}
par(mfrow = c(1, 2))
rocplot(fitted, dat[train, "y"], main = "Training Data")

svmfit.flex <- svm(y ~ ., data = dat[train, ], kernel = "radial", gamma = 50, cost = 1, decision.values = TRUE)
fitted <- attributes(predict(svmfit.flex, dat[train, ], decision.values = T))$decision.values
rocplot(fitted, dat[train, "y"], add = T, col = "red")
```

```{r}
fitted <- attributes(predict(svmfit.opt, dat[-train, ], decision.values = T))$decision.values
rocplot(fitted, dat[-train, "y"], main = "Test Data")
fitted <- attributes(predict(svmfit.flex, dat[-train, ], decision.values = T))$decision.values
rocplot(fitted, dat[-train, "y"], add = T, col = "red")
```

### 9.6.4 SVM with Multiple Classes

The [`svm()`](http://bit.ly/R_svm) function can also be used to classify observations from multiple-classes. 

```{r}
set.seed(1)
x <- rbind(x, matrix(rnorm(50 * 2), ncol = 2))
y <- c(y, rep(0, 50))
x[y == 0, 2] <- x[y == 0, 2] + 2
dat <- data.frame(x = x, y = as.factor(y))
par(mfrow = c(1, 1))
plot(x, col = (y + 1))
```

The [`svm()`](http://bit.ly/R_svm) function now will perform multi-class classification since the dataset we generated now has three class labels.

```{r}
svmfit <- svm(y ~ ., data = dat, kernel = "radial", cost = 10, gamma = 1)
plot(svmfit, dat)
```

### 9.6.5 Application to Gene Expression Data

In this final exercise we run SVM classifier on gene expression data. We load the `ISLR` package and examine the `Khan` dataset.

```{r}
library(ISLR)
names(Khan)
dim(Khan$xtrain)
dim(Khan$xtest)
length(Khan$ytrain)
length(Khan$ytest)
```

We can examine the class labels associated with the training and test subsets.

```{r}
table(Khan$ytrain)
table(Khan$ytest)
```

We use a linear kernel and run SVM classifier on the training subset.

```{r}
dat <- data.frame(x = Khan$xtrain, y = as.factor(Khan$ytrain))
out <- svm(y ~ ., data = dat, kernel = "linear", cost = 10)
summary(out)
table(out$fitted, dat$y)
```

We can then predict the classes on the test subset using the trained classifier.

```{r}
dat.te <- data.frame(x = Khan$xtest, y = as.factor(Khan$ytest))
pred.te <- predict(out, newdata = dat.te)
table(pred.te, dat.te$y)
```
